---
title: "DSC for MAD methods comparison"
author: "Kaiqian Zhang"
date: "6/11/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ggplot2)
library(gridExtra)
library(ggpubr)
```

```{r}
dscout = readRDS("data/mad-dscout.RDS")
dscout.df = data.frame(dscout$simulate.n, dscout$simulate.cp_num, dscout$simulate.residual_sd, dscout$analyze, dscout$score.error)
names(dscout.df) = c("n", "cp_num", "residual_sd", "method", "mse")
```

```{r}
plot_dscout = function(n, residual_sd){
  dscout.df1 = dscout.df[dscout.df$n==n & dscout.df$residual_sd==residual_sd,]

  dscout.df1.cp1 = dscout.df1[dscout.df1$cp_num==1,]
  cp1 = ggplot(aes(y=mse, x=method), data=dscout.df1.cp1) +
    geom_violin(alpha=0.5, color="grey") + 
    geom_jitter(alpha=0.5, aes(color=method), position = position_jitter(width = 0.1)) + 
    coord_flip() + 
    ggtitle("1 change point") + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", size=10), axis.text.x = element_text(angle = 90, hjust = 1),   axis.text.y=element_blank())
  
  dscout.df1.cp3 = dscout.df1[dscout.df1$cp_num==3,]
  cp3 = ggplot(aes(y=mse, x=method), data=dscout.df1.cp3) +
    geom_violin(alpha=0.5, color="grey") + 
    geom_jitter(alpha=0.5, aes(color=method), position = position_jitter(width = 0.1)) + 
    coord_flip() + 
    ggtitle("3 change points") + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", size=10), axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y=element_blank())
  
  dscout.df1.cp5 = dscout.df1[dscout.df1$cp_num==5,]
  cp5 = ggplot(aes(y=mse, x=method), data=dscout.df1.cp5) +
    geom_violin(alpha=0.5, color="grey") + 
    geom_jitter(alpha=0.5, aes(color=method), position = position_jitter(width = 0.1)) + 
    coord_flip() + 
    ggtitle("5 change points") + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", size=10), axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y=element_blank())
  
  dscout.df1.cp10 = dscout.df1[dscout.df1$cp_num==10,]
  cp10 = ggplot(aes(y=mse, x=method), data=dscout.df1.cp10) +
    geom_violin(alpha=0.5, color="grey") + 
    geom_jitter(alpha=0.5, aes(color=method), position = position_jitter(width = 0.1)) + 
    coord_flip() + 
    ggtitle("10 change points") + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", size=10), axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y=element_blank())
  
  dscout.df1.cp15 = dscout.df1[dscout.df1$cp_num==15,]
  cp15 = ggplot(aes(y=mse, x=method), data=dscout.df1.cp15) +
    geom_violin(alpha=0.5, color="grey") + 
    geom_jitter(alpha=0.5, aes(color=method), position = position_jitter(width = 0.1)) + 
    coord_flip() + 
    ggtitle("15 change points") + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", size=10), axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y=element_blank())
  
  dscout.df1.cp30 = dscout.df1[dscout.df1$cp_num==30,]
  cp30 = ggplot(aes(y=mse, x=method), data=dscout.df1.cp30) +
    geom_violin(alpha=0.5, color="grey") + 
    geom_jitter(alpha=0.5, aes(color=method), position = position_jitter(width = 0.1)) + 
    coord_flip() + 
    ggtitle("30 change points") + 
    theme(plot.title = element_text(hjust = 0.5, face="bold", size=10), axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y=element_blank())
  
  ggarrange(cp1, cp3, cp5, cp10, cp15, cp30, ncol=3, nrow=2, common.legend = TRUE, legend="bottom")
}
```

## Introduction

In this vignette, we compare two implemented MAD methods in `susie_trendfilter`. Our goal is to investigate whether the change made in the latter implementation has a better performance in detecting change points, which is essentially the application of `susie_trendfilter`. Two MAD implementations are illustrated below. 

- **susie_tf_MAD_original**: uses `susie_trendfilter` with orginal MAD implementation, which is 

```
estimate_mad_residual_variance_original = function(y){
  n = length(y)
  y_reflect = c(y, rev(y))
  J = floor(log2(2*n))
  y_reflect = y_reflect[1:2^J]
  y_reflect = c(y_reflect, rev(y_reflect))
  ywd = wd(y_reflect, filter.number=1, family="DaubExPhase")
  wc_d = accessD(ywd, level=J-1)
  est_resid = (median(abs(wc_d))/0.6745)^2
  return(est_resid)
}
```

- **estimate_tf_MAD_revised**: uses `susie_trendfilter` with revised MAD implementation, which is

```
estimate_mad_residual_variance_revised = function(y){
  return(0.5*(median(abs(diff(y))/0.6745)^2))
}
```
Also, we add a baseline method for comparison:

- **susie_tf**: uses `susie_trendfilter` only without MAD. 

## Simulation details

We restrict change points values between 0 and 1, and allow for the following variables:

- **n**: the number of data points. We have 100, 300, 500, 1000, or 2000 data points.

- **cp_num**: the number of change points. We have 1, 3, 5, 10, 15, 30 or 50 change points.

- **residual_sd**: residual standard deviation for the ground truth data. We allow for 0.001, 0.01, 0.02, 0.03, 0.05, and 0.1.

More details can also be seen from the [DSC script](https://github.com/stephenslab/dsc-smooth-gaussian/blob/master/susie-mad-diff.dsc). And how we simulate data can been seen from the [R code (function `simulate_tf_order0`)](https://github.com/stephenslab/dsc-smooth-gaussian/blob/master/code/simulate.R).

## Results summary

- Given 100 data points, the revised implementation overall has a better performance than the original one.

- Given 1000 data points, we cannot observe much difference between the two implementations.

- The revised MAD implementation, by DSC results, should replace the original one because of simplicity and improved performance.

## Results details

### Given 100 data points

Given 100 data points and assume residual standard deviation is 0.001.
```{r}
plot_dscout(n=100, residual_sd=0.001)
```

Given 100 data points and assume residual standard deviation is 0.01.
```{r}
plot_dscout(n=100, residual_sd=0.01)
```

Given 100 data points and assume residual standard deviation is 0.05.
```{r}
plot_dscout(n=100, residual_sd=0.05)
```

Given 100 data points and assume residual standard deviation is 0.1.
```{r}
plot_dscout(n=100, residual_sd=0.1)
```

### Given 1000 data points

Given 1000 data points and assume residual standard deviation is 0.001.
```{r}
plot_dscout(n=1000, residual_sd=0.001)
```

Given 1000 data points and assume residual standard deviation is 0.01.
```{r}
plot_dscout(n=1000, residual_sd=0.01)
```

Given 1000 data points and assume residual standard deviation is 0.05.
```{r}
plot_dscout(n=1000, residual_sd=0.05)
```

Given 1000 data points and assume residual standard deviation is 0.1.
```{r}
plot_dscout(n=1000, residual_sd=0.1)
```












