---
title: "susie-residual-20190315"
author: "Kaiqian Zhang"
date: "3/15/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---

Here we use a robust estimator MAD to estimate residual variance. Donoho & Johnstone (1994) considered estimating $\sigma$ in the wavelet domain and suggested a robust estimate that is based only on the empirical wavelet coefficients at the finest resolution level. The reason for considering only the finest level is that corresponing empirical wavelet
coefficients tend to consist mostly of noise. Donoho & Johnstone (1994) proposed a robust estimate of the noise level $\sigma$ (based on the median absolute deviation) given by

$$
\hat{\sigma} = \frac{\text{median}(\{|\hat{d}_{J-1,k}|: k=0,1,\dots,2^{J-1}-1\})}{0.6745},
$$
where $\hat{d}_{J-1,k}$ are mother wavelet coefficients at the finest level and $\hat{\sigma}$ is the estimated residual standard deviation.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(susieR)
library(wavethresh)
library(binhf) #Package for shift function
library(smashr)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  #m = 10% * abs(x_i - x_{i+1})
  m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
  K = 2^(ceiling(log2(m)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}

#' @param K is the dimension of decinmated W
#' @return a K by K matrix, which the transpose of decimated Harr wavelet transform matrix 
create_Wt_DWT = function(K){
  return(GenW(n=K, filter.number=1, family="DaubExPhase"))
}

create_y = function(x, beta.function, n, sd, boundaries, beta_vals){
  set.seed(1)
  return(beta.function(x, boundaries, beta_vals) + rnorm(n, sd=sd))
}
```


## piecewise-constant with three intervals example

```{r}
interval_num = 3
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta3 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else{
      beta[i] = beta_vals[3]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta3, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta3(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```


## piecewise-constant with four intervals example
```{r}
interval_num = 4
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta4 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else{
      beta[i] = beta_vals[4]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta4, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta4(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```


## piecewise-constant with five intervals example

```{r}
interval_num = 5
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta5 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else{
      beta[i] = beta_vals[5]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta5, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta5(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```

## piecewise-constant with six intervals example

```{r}
interval_num = 6
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta6 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else if(x[i] >= boundaries[4] & x[i] < boundaries[5]){
      beta[i] = beta_vals[5]
    } else{
      beta[i] = beta_vals[6]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta6, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta6(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```

## pieces-constant with seven intervals example

```{r}
interval_num = 7
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta7 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else if(x[i] >= boundaries[4] & x[i] < boundaries[5]){
      beta[i] = beta_vals[5]
    } else if(x[i] >= boundaries[5] & x[i] < boundaries[6]){
      beta[i] = beta_vals[6]
    } else{
      beta[i] = beta_vals[7]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta7, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta7(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```
## pieces-constant with ten intervals example

```{r}
interval_num = 10
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta10 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else if(x[i] >= boundaries[4] & x[i] < boundaries[5]){
      beta[i] = beta_vals[5]
    } else if(x[i] >= boundaries[5] & x[i] < boundaries[6]){
      beta[i] = beta_vals[6]
    } else if(x[i] >= boundaries[6] & x[i] < boundaries[7]){
      beta[i] = beta_vals[7]
    } else if(x[i] >= boundaries[7] & x[i] < boundaries[8]){
      beta[i] = beta_vals[8]
    } else if(x[i] >= boundaries[8] & x[i] < boundaries[9]){
      beta[i] = beta_vals[9]
    } else{
      beta[i] = beta_vals[10]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta10, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta10(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```
