---
title: "susie-symlet-20190319"
author: "Kaiqian Zhang"
date: "3/19/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---

Previously we investigated estimating piecewise-constant functions by using Haar wavelets. In this vignette we try to use Symlet wavelets to estimate various functions, such as spikes, bumps, blocks, angles, doppler, and blip function as provided in the dsc-smooth-gaussian repo.

The following figures show SuSiE fits for simulated data:

* black line: underlying truth.

* orange line: Symlet wavelts + SuSiE fit initialized from a fit with very small and fixed residual variance.

* blue line: Symlet wavelets + SuSiE fit with fixed residual variance estimated using MAD.

* brown line: Haar wavelets + SuSiE fit with fixed residual variance estimated using MAD.

## Results summary

* We obtain good fits in esitmating doppler, spikes, and bumps functions by using Symlet wavelets.

* Blocks function, i.e. piecewise-constant function is better to use Haar wavelets to estimate. Using Symlet wavelets is not a good idea here.

* Angles and blip functions do not achieve a nice fit either considering Haar or Symlet wavelets. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(susieR)
library(wavethresh)
library(binhf) #Package for shift function
library(smashr)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  #m = 10% * abs(x_i - x_{i+1})
  m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
  K = 2^(ceiling(log2(m)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}

#' @param K is the dimension of decinmated W
#' @return a K by K matrix, which the transpose of decimated Harr wavelet transform matrix 
create_Wt_DWT = function(K, type){
  if (type=="Haar"){
    return(GenW(n=K, filter.number=1, family="DaubExPhase"))
  } else if (type=="Symlet"){
    return(GenW(n=K, filter.number=10, family="DaubLeAsymm", bc="periodic"))
  }else{
    stop("Input type should be either Haar or Symlet.")
  }
}

create_y = function(x, beta.function, n, sd, boundaries, beta_vals){
  set.seed(1)
  return(beta.function(x, boundaries, beta_vals) + rnorm(n, sd=sd))
}
```

```{r}
spikes.fn = function(t) {
  (0.75 * exp(-500 * (t - 0.23)^2) + 1.5 * exp(-2000 * (t - 0.33)^2) + 3 * exp(-8000 * (t - 0.47)^2) + 2.25 * exp(-16000 * (t - 0.69)^2) + 0.5 * exp(-32000 * (t - 0.83)^2))
}

bumps.fn = function(t) {
  pos = c(0.1, 0.13, 0.15, 0.23, 0.25, 0.4, 0.44, 0.65, 0.76, 0.78, 0.81)
  hgt = 2.97/5 * c(4, 5, 3, 4, 5, 4.2, 2.1, 4.3, 3.1, 5.1, 4.2)
  wth = c(0.005, 0.005, 0.006, 0.01, 0.01, 0.03, 0.01, 0.01, 0.005, 0.008, 0.005)
  fn = rep(0, length(t))
  for (j in 1:length(pos)) {
    fn = fn + hgt[j]/((1 + (abs(t - pos[j])/wth[j]))^4)
  }
  return((1 + fn)/5)
}

blocks.fn = function(t) {
  pos = c(0.1, 0.13, 0.15, 0.23, 0.25, 0.4, 0.44, 0.65, 0.76, 0.78, 0.81)
  hgt = 2.88/5 * c(4, (-5), 3, (-4), 5, (-4.2), 2.1, 4.3, (-3.1), 2.1, (-4.2))
  fn = rep(0, length(t))
  for (j in 1:length(pos)) {
    fn = fn + (1 + sign(t - pos[j])) * (hgt[j]/2)
  }
  return(0.2 + 0.6 * (fn - min(fn))/max(fn - min(fn)))
}

angles.fn = function(t) {
  sig = ((2 * t + 0.5) * (t <= 0.15)) + ((-12 * (t - 0.15) + 0.8) * (t > 0.15 & t <= 0.2)) + 0.2 * (t > 0.2 & t <= 0.5) + ((6 * (t - 0.5) + 0.2) * (t > 0.5 & t <= 0.6)) + ((-10 * (t - 0.6) + 0.8) * (t > 0.6 & t <= 0.65)) + ((-0.5 * (t - 0.65) + 0.3) * (t > 0.65 & t <= 0.85)) + ((2 * (t - 0.85) + 0.2) * (t > 0.85))
  fn = 3/5 * ((5/(max(sig) - min(sig))) * sig - 1.6) - 0.0419569
  return((1 + fn)/5)
}

doppler.fn = function(t) {
  dop.f = function(x) sqrt(x * (1 - x)) * sin((2 * pi * 1.05)/(x + 0.05))
  fn = dop.f(t)
  fn = 3/(max(fn) - min(fn)) * (fn - min(fn))
  return((1 + fn)/5)
}

blip.fn = function(t) {
  fn = (0.32 + 0.6 * t + 0.3 * exp(-100 * (t - 0.3)^2)) * (t >= 0 & t <= 0.8) + (-0.28 + 0.6 * t + 0.3 * exp(-100 * (t - 1.3)^2)) * (t > 0.8 & t <= 1)
  return(fn)
}
```

## Symlet wavelets on different functions

### Doppler function

```{r}
sd = 0.01
n = 200
set.seed(1)
x = sort(runif(n, 0, 1))
y = doppler.fn(x) + rnorm(n ,0, sd=sd)
```

```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Symlet")
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=10, family="DaubLeAsymm", type="wavelet",
bc="periodic", verbose=FALSE, min.scale=0, precond=TRUE)
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, doppler.fn(x), col="black", lwd=2)
lines(x, predict(s.initfix), col="orange", lwd=2)
lines(x, predict(s.est_resid), col="blue", lwd=2)
```

### Spikes function

```{r}
sd = 0.01
n = 200
set.seed(1)
x = sort(runif(n, 0, 1))
y = spikes.fn(x) + rnorm(n ,0, sd=sd)
```

```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Symlet")
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=10, family="DaubLeAsymm", type="wavelet",
bc="periodic", verbose=FALSE, min.scale=0, precond=TRUE)
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, spikes.fn(x), col="black", lwd=2)
lines(x, predict(s.initfix), col="orange", lwd=2)
lines(x, predict(s.est_resid), col="blue", lwd=2)
```

### Bumps function

```{r}
sd = 0.01
n = 200
set.seed(1)
x = sort(runif(n, 0, 1))
y = bumps.fn(x) + rnorm(n ,0, sd=sd)
```

```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Symlet")
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=10, family="DaubLeAsymm", type="wavelet",
bc="periodic", verbose=FALSE, min.scale=0, precond=TRUE)
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, bumps.fn(x), col="black", lwd=2)
lines(x, predict(s.initfix), col="orange", lwd=2)
lines(x, predict(s.est_resid), col="blue", lwd=2)
```


### Blocks function (i.e. piecewise-constant function)
```{r}
sd = 0.01
n = 200
set.seed(1)
x = sort(runif(n, 0, 1))
y = blocks.fn(x) + rnorm(n ,0, sd=sd)
```

```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Symlet")
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=10, family="DaubLeAsymm", type="wavelet",
bc="periodic", verbose=FALSE, min.scale=0, precond=TRUE)
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```


```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Haar")
RWt = R %*% Wt.DWT
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid.Haar = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, blocks.fn(x), col="black", lwd=2)
lines(x, predict(s.initfix), col="orange", lwd=2)
lines(x, predict(s.est_resid), col="blue", lwd=2)
lines(x, predict(s.est_resid.Haar), col="brown", lwd=2)
```

### Angles function

```{r}
sd = 0.01
n = 200
set.seed(1)
x = sort(runif(n, 0, 1))
y = angles.fn(x) + rnorm(n ,0, sd=sd)
```

```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Symlet")
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=10, family="DaubLeAsymm", type="wavelet",
bc="periodic", verbose=FALSE, min.scale=0, precond=TRUE)
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```
```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Haar")
RWt = R %*% Wt.DWT
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=1, family="DaubExPhase")
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid.Haar = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, angles.fn(x), col="black", lwd=2)
lines(x, predict(s.initfix), col="orange", lwd=2)
lines(x, predict(s.est_resid), col="blue", lwd=2)
lines(x, predict(s.est_resid.Haar), col="brown", lwd=2)
```

### Blip function

```{r}
sd = 0.01
n = 200
set.seed(1)
x = sort(runif(n, 0, 1))
y = blip.fn(x) + rnorm(n ,0, sd=sd)
```

```{r}
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K, "Symlet")
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
y_reflect = c(y, rev(y))
J = floor(log2(2*n))
y_reflect = y_reflect[1:2^J]
y_reflect = c(y_reflect, rev(y_reflect))
ywd <- wd(y_reflect, filter.number=10, family="DaubLeAsymm", type="wavelet",
bc="periodic", verbose=FALSE, min.scale=0, precond=TRUE)
wc_d = accessD(ywd, level=J-1)
est.resid = (median(abs(wc_d))/0.6745)^2
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, blip.fn(x), col="black", lwd=2)
lines(x, predict(s.initfix), col="orange", lwd=2)
lines(x, predict(s.est_resid), col="blue", lwd=2)
```























