---
title: "susie-residual-20190313"
author: "Kaiqian Zhang"
date: "3/13/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---

Previously, we notice that when setting the residual variance as a very small fixed value, we could obtain a nice fit in SuSiE. We, however, cannot simply do this trick since SuSiE will do everything it can to reduce variance and start to fit noise as signal. We can have the following two options:

(1) initialize from the fit with a small residual variance. [red line in the figure]

(2) use `smashr::sd_estimate_gasser_etal` function to estimate residual variance. This estimator is from Equation(3) in Brown and Levine (2007), which uses three neighboring points to estimate the residual variance. [green line in the figure]

(3) use the fit from (2) as an initialization in SuSiE. [orange line in the figure]

We simulate piecewise-constant nonparametric regression examples to see the performance.

Results: Both (1) and (3) can give good fits, but (3) is quicker and has no "IBSS algorithm did not converge in 100 iterations!" warning issue. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(susieR)
library(wavethresh)
library(binhf) #Package for shift function
library(smashr)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  #m = 10% * abs(x_i - x_{i+1})
  m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
  K = 2^(ceiling(log2(m)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}

#' @param K is the dimension of decinmated W
#' @return a K by K matrix, which the transpose of decimated Harr wavelet transform matrix 
create_Wt_DWT = function(K){
  return(GenW(n=K, filter.number=1, family="DaubExPhase"))
}

create_y = function(x, beta.function, n, sd, boundaries, beta_vals){
  set.seed(1)
  return(beta.function(x, boundaries, beta_vals) + rnorm(n, sd=sd))
}
```


## piecewise-constant with three intervals example

```{r}
interval_num = 3
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta3 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else{
      beta[i] = beta_vals[3]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta3, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
est.resid = sd_estimate_gasser_etal(y)
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta3(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```


## piecewise-constant with four intervals example
```{r}
interval_num = 4
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta4 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else{
      beta[i] = beta_vals[4]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta4, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
est.resid = sd_estimate_gasser_etal(y)
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta4(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```


## piecewise-constant with five intervals example

```{r}
interval_num = 5
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta5 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else{
      beta[i] = beta_vals[5]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta5, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
est.resid = sd_estimate_gasser_etal(y)
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta5(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```

## piecewise-constant with six intervals example

```{r}
interval_num = 6
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta6 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else if(x[i] >= boundaries[4] & x[i] < boundaries[5]){
      beta[i] = beta_vals[5]
    } else{
      beta[i] = beta_vals[6]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta6, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
est.resid = sd_estimate_gasser_etal(y)
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta6(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```

## pieces-constant with seven intervals example

```{r}
interval_num = 7
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta7 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else if(x[i] >= boundaries[4] & x[i] < boundaries[5]){
      beta[i] = beta_vals[5]
    } else if(x[i] >= boundaries[5] & x[i] < boundaries[6]){
      beta[i] = beta_vals[6]
    } else{
      beta[i] = beta_vals[7]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta7, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
est.resid = sd_estimate_gasser_etal(y)
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta7(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```
## pieces-constant with ten intervals example

```{r}
interval_num = 10
boundaries = round(sort(runif(interval_num-1)),2)
beta_vals = sample(seq(-10,10), interval_num)
```

```{r}
f.beta10 = function(x, boundaries, beta_vals){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < boundaries[1]){
      beta[i] = beta_vals[1]
    } else if(x[i] >= boundaries[1] & x[i] < boundaries[2]){
      beta[i] = beta_vals[2]
    } else if(x[i] >= boundaries[2] & x[i] < boundaries[3]){
      beta[i] = beta_vals[3]
    } else if(x[i] >= boundaries[3] & x[i] < boundaries[4]){
      beta[i] = beta_vals[4]
    } else if(x[i] >= boundaries[4] & x[i] < boundaries[5]){
      beta[i] = beta_vals[5]
    } else if(x[i] >= boundaries[5] & x[i] < boundaries[6]){
      beta[i] = beta_vals[6]
    } else if(x[i] >= boundaries[6] & x[i] < boundaries[7]){
      beta[i] = beta_vals[7]
    } else if(x[i] >= boundaries[7] & x[i] < boundaries[8]){
      beta[i] = beta_vals[8]
    } else if(x[i] >= boundaries[8] & x[i] < boundaries[9]){
      beta[i] = beta_vals[9]
    } else{
      beta[i] = beta_vals[10]
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta10, n, sd, boundaries, beta_vals)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
```

```{r}
s.fix = susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = 0.01)
```

```{r}
s.initfix = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.fix)
```

```{r}
est.resid = sd_estimate_gasser_etal(y)
```

```{r}
s.est_resid= susie(RWt, y, L=50, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', estimate_residual_variance = FALSE, residual_variance = est.resid)
```

```{r}
s.initest_resid = susie(RWt, y, estimate_prior_variance = TRUE, estimate_prior_method = 'optim', s_init = s.est_resid)
```


```{r}
plot(x, y, pch=20, col="grey")
lines(x, f.beta10(x, boundaries, beta_vals), col="black", lwd=2)
lines(x, predict(s.initfix), col="red", lwd=2)
lines(x, predict(s.est_resid), col="green", lwd=2)
lines(x, predict(s.initest_resid), col="orange", lwd=2)
```
