---
title: "Undecimated Wavelet Transform"
author: "Kaiqian Zhang"
date: "1/14/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In this vignette, we explore undecimated wavelet transform for sparse regression. We simulate piecewise constant data `y` which has `n=2^7=128` samples. We use two different bases:

1) Discrete wavelet transform (DWT): each basis are orthonormal to each other;

2) Undecimated wavelet transform (UDWT): we have n*log2(n) bases and they are not orthogonal.

We expect that the fit by UDWT will be smoother than that by DWT. 

```{r, warning=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
```

## Create DWT and UDWT matrix
```{r}
J = 7
n = 2^J
```

```{r}
Haar_matrix <- t(GenW(n=n, filter.number=1, family="DaubExPhase"))
```

```{r}
Haar_matrix_inv <- t(Haar_matrix)
dim(Haar_matrix_inv)
```

```{r}
n_Haar_matrix_inv <- Haar_matrix_inv
for(i in 1:(n-1)){
  order <- shift(seq(n),i)
  Haar_matrix_shifted_inv <- t(Haar_matrix[,order])
  n_Haar_matrix_inv <- cbind(n_Haar_matrix_inv, Haar_matrix_shifted_inv) #Each column is a basis for sparse regression
}
```

```{r}
duplicated_columns <- duplicated(t(n_Haar_matrix_inv))
W_T_Haar_nlog2n <- n_Haar_matrix_inv[, !duplicated_columns]
```

```{r}
dim(W_T_Haar_nlog2n)
```

Notice that matrix `X` used in DWT is `r n` by `r n` (`Haar_matrix_inv`) and matrix X used in UDWT is `r dim(W_T_Haar_nlog2n)[1]` by `r dim(W_T_Haar_nlog2n)[2]`. (`W_T_Haar_nlog2n`)

## Simulate data `y`

```{r}
set.seed(1)
beta <- c(rep(0,10),rep(1,20),rep(-2,10),rep(0,n-40))
y <- beta + rnorm(n, sd=0.1)
```

## Fit SuSiE with DWT bases
```{r}
fit_susie_Haar <- susie(Haar_matrix_inv, y, L=100)
```

```{r}
plot(y,pch=20, col="grey", main="SuSiE-DWT")
lines(beta,col="black",lwd=2)
lines(predict(fit_susie_Haar),col="red",lwd=2)
```

## Fit SuSiE with UDWT bases
```{r}
fit_susie_Haar_nlog2n <- susie(W_T_Haar_nlog2n, y, L=100)
```

```{r}
plot(y,pch=20, col="grey", main="SuSiE-UDWT")
lines(beta,col="black",lwd=2)
lines(predict(fit_susie_Haar_nlog2n),col="red",lwd=2)
```

If we further estimate prior variance in SuSiE,

```{r}
fit_susie_Haar_nlog2n_prior <- susie(W_T_Haar_nlog2n, y, L=100, estimate_prior_variance = TRUE)
```

```{r}
plot(y,pch=20, col="grey", main="SuSiE-UDWT with prior variance estimated")
lines(beta,col="black",lwd=2)
#lines(predict(fit_susie_Haar_nlog2n),col="red",lwd=2)
lines(predict(fit_susie_Haar_nlog2n_prior), col='blue', lwd=2)
```



```{r}
library(smashr)
smashr_y = smash.gaus(y)
```


```{r}
plot(y,pch=20, col="grey", main="smashr")
lines(beta,col="black",lwd=2)
lines(smashr_y,col="red",lwd=2)
```

## What is going wrong? Try initialization

We observe that the performance of `SuSiE-DWT` is not decent but `SuSiE-trend filering` has a great performance. And we want to understand whether the bad performance of `SuSiE with UDWT` is due to convergence of local optimum. So we initialize SuSiE from its trend filtering fit and then fit the `SuSiE with UDWT` model. But the result is not in my expectation (I expect a good fit due to initialization) and what other steps I could take to understand why?
```{r}
fit_susie_trendfilter<- susie_trendfilter(y, order=0)
```

```{r}
plot(y,pch=20, col="grey", main="SuSiE-trend filtering")
lines(beta,col="black",lwd=2)
lines(predict(fit_susie_trendfilter),col="red",lwd=2)
```

```{r}
coef_index <- seq(1,127)
coef_value <- coef(fit_susie_trendfilter)[2:128]
p <- dim(W_T_Haar_nlog2n)[2]
susie_init_trendfilter = susie_init_coef(coef_index, coef_value, p)
```

```{r}
fit_susie_UDWT_init <- susie(W_T_Haar_nlog2n, y, s_init=susie_init_trendfilter)
```

```{r}
plot(y,pch=20, col="grey", main="SuSiE-UWDT with trend filtering initialization")
lines(beta,col="black",lwd=2)
lines(predict(fit_susie_UDWT_init),col="red",lwd=2)
```

















