---
title: "Undecimated Wavelet Transform"
author: "Kaiqian Zhang"
date: "1/14/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
```
In this vignette, we apply undecimated wavelet bases on our sparse regression method `SuSiE` and try to investigate the performance. We compare two different bases:

1) Discrete wavelet transform (DWT): each basis are orthonormal to each other;

2) Undecimated wavelet transform (UDWT): we have n*log2(n) bases and they are not orthogonal.

We expect that fit by UDWT will be smoother than that by DWT. 

## Last time investigation
Last time, we simulated piecewise constant data of size `n = 512`. We discovered that using discrete wavelet bases (DWT) actually is better than using redundant ones (UDWT). We can observe this by red lines in the figures `SuSiE-DWT n=512` and `SuSiE-UDWT n=512`.  

We hypothesize that the worse fit for UDWT might due to convergence to local optimum problem. So we initialize `SuSiE` from DWT fit, using coefficients that have absolute value greater than 0.01. Then we fit the model with UDWT bases using our initialization. Please see green line in the figure `SuSiE-UDWT n=512` below. The fit is not very good. So my questions are:

1) Does the performance of green line suggest the convergence to local optimum problem? I expect that with DWT initialization, the fit will be more like DWT fit. But it turns out not.

2) Blue lines below indicate we estimate prior variance when fitting SuSiE. I assume estimating prior variance would always improve the fit. But in the figure `SuSiE-DWT n=512`, estimating prior variance on the other hand makes the fit worse. 

### Simulate data `n=512`
We simulate piecewise constant data of size `n=512`.
```{r}
n <- 512
set.seed(1)
beta1 <- c(rep(0,100),rep(1,100),rep(3,100),rep(-2,100),rep(0,n-400))
y1 <- beta1 + rnorm(n, sd=0.1)
```

### Create DWT and UDWT matrices as input `X`
We create matrices using DWT and UDWT bases when `n=512`. 
```{r}
Haar_matrix <- t(GenW(n=n, filter.number=1, family="DaubExPhase"))
DWT_W_T <- t(Haar_matrix)
n_Haar_matrix_inv <- DWT_W_T
for(i in 1:(n-1)){
  order <- shift(seq(n),i)
  Haar_matrix_shifted_inv <- t(Haar_matrix[,order])
  n_Haar_matrix_inv <- cbind(n_Haar_matrix_inv, Haar_matrix_shifted_inv) #Each column is a basis for sparse regression
}
duplicated_columns <- duplicated(t(n_Haar_matrix_inv))
UDWT_W_T <- n_Haar_matrix_inv[, !duplicated_columns]
```

### Fit SuSiE with DWT bases
We fit SuSiE with DWT bases when `n=512`. The red line is SuSiE fit, and the blue line is SuSiE fit with `estimate_prior_variance=TRUE`. 
```{r}
fit_susie_DWT = susie(DWT_W_T, y1, L=30)
```

```{r}
fit_susie_DWT_prior = susie(DWT_W_T, y1, L=50, estimate_prior_variance=TRUE)
```

```{r}
plot(y1,pch=20, col="grey", main="SuSiE-DWT n=512")
lines(beta1,col="black",lwd=2)
lines(predict(fit_susie_DWT), col='red', lwd=2)
lines(predict(fit_susie_DWT_prior), col='blue', lwd=2)
```

### Fit SuSiE with UDWT bases
We fit SuSiE with UDWT bases when `n=512`. The red line is SuSiE fit, the blue line is SuSiE fit with `estimate_prior_variance=TRUE`, and the green line is SuSiE fit with initialization from DWT fit.

```{r}
fit_susie_UDWT = susie(UDWT_W_T, y1, L=30)
```

```{r}
fit_susie_UDWT_prior = susie(UDWT_W_T, y1, L=30, estimate_prior_variance=TRUE)
```

```{r}
coef_index <- which(abs(coef(fit_susie_DWT))>0.01)
coef_value <- coef(fit_susie_DWT)[coef_index]
coef_index <- coef_index[2:length(coef_index)]-1
coef_value <- coef_value[2:length(coef_value)]
p <- dim(UDWT_W_T)[2]
susie_initDWT <- susie_init_coef(coef_index, coef_value, p)
```

```{r}
fit_susie_UDWT_init <- susie(UDWT_W_T, y1, s_init = susie_initDWT)
```

```{r}
plot(y1,pch=20, col="grey", main="SuSiE-UDWT n=512")
lines(beta1,col="black",lwd=2)
lines(predict(fit_susie_UDWT), col='red', lwd=2)
lines(predict(fit_susie_UDWT_prior), col='blue', lwd=2)
lines(predict(fit_susie_UDWT_init), col='green', lwd=2)
```

## Investigation with smaller sample size
To understand what is going on, we simplify the sample size to `n=128`. When my sample size becomes `n=128`, I notice that using UDWT and estimating prior variance would be better than using DWT. But I still do not know why it is worse when estimating prior variance in the DWT case. 

### Simulate data `n=128`
```{r}
n <- 128
set.seed(1)
beta2 <- c(rep(0,50),rep(-2,50),rep(0,n-100))
y2 <- beta2 + rnorm(n, sd=0.1)
```

### Create DWT and UDWT matrices as input `X`
We create matrices using DWT and UDWT bases when `n=128`. 
```{r}
Haar_matrix <- t(GenW(n=n, filter.number=1, family="DaubExPhase"))
DWT_W_T <- t(Haar_matrix)
n_Haar_matrix_inv <- DWT_W_T
for(i in 1:(n-1)){
  order <- shift(seq(n),i)
  Haar_matrix_shifted_inv <- t(Haar_matrix[,order])
  n_Haar_matrix_inv <- cbind(n_Haar_matrix_inv, Haar_matrix_shifted_inv) #Each column is a basis for sparse regression
}
duplicated_columns <- duplicated(t(n_Haar_matrix_inv))
UDWT_W_T <- n_Haar_matrix_inv[, !duplicated_columns]
```

### Fit SuSiE with DWT bases
We fit SuSiE with DWT bases when `n=128`. The red line is SuSiE fit, and the blue line is SuSiE fit with `estimate_prior_variance=TRUE`. 
```{r}
fit_susie_DWT = susie(DWT_W_T, y2, L=30)
```

```{r}
fit_susie_DWT_prior = susie(DWT_W_T, y2, L=50, estimate_prior_variance=TRUE)
```

```{r}
plot(y2,pch=20, col="grey", main="SuSiE-DWT n=128")
lines(beta2,col="black",lwd=2)
lines(predict(fit_susie_DWT), col='red', lwd=2)
lines(predict(fit_susie_DWT_prior), col='blue', lwd=2)
```


### Fit SuSiE with UDWT bases
We fit SuSiE with UDWT bases when `n=128`. The red line is SuSiE fit, the blue line is SuSiE fit with `estimate_prior_variance=TRUE`.

```{r}
fit_susie_UDWT = susie(UDWT_W_T, y2, L=30)
```

```{r}
fit_susie_UDWT_prior = susie(UDWT_W_T, y2, L=30, estimate_prior_variance=TRUE)
```

```{r}
plot(y2,pch=20, col="grey", main="SuSiE-UDWT n=128")
lines(beta2,col="black",lwd=2)
lines(predict(fit_susie_UDWT), col='red', lwd=2)
lines(predict(fit_susie_UDWT_prior), col='blue', lwd=2)
```





