---
title: "SuSiE Estimate Prior Variance Methods Comparison (DSC)"
author: "Kaiqian Zhang"
date: "2/13/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

We use datasets from `dsc-finemap` and perform DSC simulations to compare three methods `uniroot`, `optim`, and `EM`. Implementations for those three methods are:

* uniroot: 

```
V.u=uniroot(negloglik.grad.logscale,c(-10,10),extendInt = "upX",betahat=betahat,shat2=shat2,prior_weights=prior_weights)
V = exp(V.u$root)
if(loglik(0,betahat,shat2,prior_weights) >= loglik(V,betahat,shat2,prior_weights)){
      V=0 
  }
```

* optim:

```
lV = optim(par=log(max(betahat^2-shat2, na.rm = TRUE)), fn=neg.loglik.logscale, gr=negloglik.grad.logscale, betahat=betahat, shat2=shat2, prior_weights = prior_weights, method='Brent', lower = -10, upper = 10)$par
V = exp(lV)
if(loglik(0,betahat,shat2,prior_weights) >= loglik(V,betahat,shat2,prior_weights)){
    V=0 
  }
```

* EM:

```
V = sum(alpha*post_mean2)
```

## Evaluation criteria

* converged: We perform two replicates in the simulation. If converged is 2, SuSiE converges.

* objective: the objective of SuSiE fit. Since we have two replicates, we have objective1 and objective2 in our results. We expect that the higher the objective, the better the method. 

## Results

*Summary:*

- In this simulation study, all methods converge.

- Objectives produced are very similar. But `optim` consistently and slightly outperforms the other two.

## Tables and code details

```{r}
dscout = readRDS("/Users/Kaiqian /Desktop/wavelet/susie-np/data/dsc_em_output.RDS")
```

```{r}
dscout.df = data.frame(dscout$method,dscout$simple_lm.n_signal, dscout$simple_lm.pve, dscout$score_susie.total, dscout$score_susie.valid, dscout$score_susie.size, dscout$score_susie.purity, dscout$score_susie.top, dscout$score_susie.objective.1, dscout$score_susie.objective.2, dscout$score_susie.converged)
names(dscout.df) = c("method", "n_signal", "pve", "total", "valid", "size", "purity", "top", "objective1", "objective2", "converged")
```

We have the following number of signals in the simulation.
```{r}
unique(dscout$simple_lm.n_signal)
```

We have the following options for pve in the simulation.
```{r}
unique(dscout$simple_lm.pve)
```

When the number of signals is 1, two objectives from two replicates are computed. 
```{r}
dscout.df1 = dscout.df[dscout.df$n_signal==1,]
dscout.df.n1o1.summary = aggregate(objective1~method+pve+converged, dscout.df1, mean)
dscout.df.n1o2.summary = aggregate(objective2~method+pve+converged, dscout.df1, mean)
dscout.df.n1o1.summary
dscout.df.n1o2.summary
```

When the number of signals is 2, two objectives from two replicates are computed. 
```{r}
dscout.df2 = dscout.df[dscout.df$n_signal==2,]
dscout.df.n2o1.summary = aggregate(objective1~method+pve+converged, dscout.df2, mean)
dscout.df.n2o2.summary = aggregate(objective2~method+pve+converged, dscout.df2, mean)
dscout.df.n2o1.summary
dscout.df.n2o2.summary
```

When the number of signals is 3, two objectives from two replicates are computed. 
```{r}
dscout.df3 = dscout.df[dscout.df$n_signal==3,]
dscout.df.n3o1.summary = aggregate(objective1~method+pve+converged, dscout.df3, mean)
dscout.df.n3o2.summary = aggregate(objective2~method+pve+converged, dscout.df3, mean)
dscout.df.n3o1.summary
dscout.df.n3o2.summary
```

When the number of signals is 4, two objectives from two replicates are computed. 
```{r}
dscout.df4 = dscout.df[dscout.df$n_signal==4,]
dscout.df.n4o1.summary = aggregate(objective1~method+pve+converged, dscout.df4, mean)
dscout.df.n4o2.summary = aggregate(objective2~method+pve+converged, dscout.df4, mean)
dscout.df.n4o1.summary
dscout.df.n4o2.summary
```

When the number of signals is 5, two objectives from two replicates are computed. 
```{r}
dscout.df5 = dscout.df[dscout.df$n_signal==5,]
dscout.df.n5o1.summary = aggregate(objective1~method+pve+converged, dscout.df5, mean)
dscout.df.n5o2.summary = aggregate(objective2~method+pve+converged, dscout.df5, mean)
dscout.df.n5o1.summary
dscout.df.n5o2.summary
```
