---
title: "Wavelet-based SuSiE Analysis 20190228"
author: "Kaiqian Zhang"
date: "2/28/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this vignette, I will show two very similar simulation examples. But one has a really good performance but the other does not. I initially had a really bad fit for a more complicated version of simulation example 2. I have more change points in my orginal simulation example 2. After observing that, I simplified my example to the current simulation example 2. 

And for simulation example 1, I notice that even if I add more change points on top of it, I can still have a good fit. This is strange to me since I feel like those two simulations are very similar: they both have abrupt jumps within a window of 0.05, but their fit performances are totally different. 

## Set up and functions

```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
```

Function for creating interpolation matrix `R`:
```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

Function for creating decimated Wt:
```{r}
#' @param K is the dimension of decinmated W
#' @return a K by K matrix, which the transpose of decimated Harr wavelet transform matrix 
create_Wt_DWT = function(K){
  return(GenW(n=K, filter.number=1, family="DaubExPhase"))
}
```

```{r}
create_y = function(x, beta.function, n, sd){
  set.seed(1)
  return(beta.function(x) + rnorm(n, sd=sd))
}
```

## Simulation example 1

```{r}
f.beta3 = function(x){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < 0.5){
      beta[i] = 0
    } else if(x[i] >= 0.5 & x[i] < 0.55){
      beta[i] = 7
    } else if(x[i] >= 0.55 & x[i] < 0.6){
      beta[i] = -4
    } else{
      beta[i] = 0
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta3, n, sd)
```

```{r}
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
#Run SuSiE
s = susie(RWt, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 1: fit for simulation example 1")
lines(x, f.beta3(x), col="black", lwd=2)
lines(x, predict(s), col="blue", lwd=2)
```

## Simulation example 2

```{r}
f.beta4 = function(x){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < 0.6){
      beta[i] = -3
    } else if(x[i] >= 0.6 & x[i] < 0.65){
      beta[i] = 7
    } else if(x[i] >= 0.65 & x[i] < 0.7){
      beta[i] = 0
    } else {
      beta[i] = 2
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta4, n, sd)
```

```{r}
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
#Run SuSiE
s = susie(RWt, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 2: fit for simulation example 2")
lines(x, f.beta4(x), col="black", lwd=2)
lines(x, predict(s), col="blue", lwd=2)
```
