---
title: "Wavelet-based SuSiE Analysis 20190127"
author: "Kaiqian Zhang"
date: "1/27/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
library(glmnet)
library(smashr)
```

In this analysis, we use DWT (i.e. orthogonal wavelet basis) in the SuSiE. Our goal is to apply wavelet-based method on non-equispaced data with sample size not necessarily a power of 2. We try a new interpolation approach proposed by Asad et al. in 2018 [(paper link)](https://papers.nips.cc/paper/8112-wavelet-regression-and-additive-models-for-irregularly-spaced-data.pdf). 

## wavelet-based SuSiE problem

A non-parametric regression problem in the wavelet context is 

$$
\hat{b} = \underset{b}{\mathrm{argmin}} \frac{1}{2}|| y- b||_2^2 + \lambda||Wb||_1,
$$
where $W$ is the discrete wavelet transform (UWD) matrix. 

Let $\beta=Wb$. Then $b = W^T\beta$ since $W$ is orthogonal. The above problem becomes

$$\hat{\beta} = \underset{\beta}{\mathrm{argmin}} \frac{1}{2}||y- W^T\beta||_2^2 + \lambda||\beta||_1.$$

## Asad's interpolation matrix

Given that $y$ has $n$ data points, $W$ is an $n$ by $n$ matrix. The DWT, however, requires $n$ to be a power of 2. We here try a novel interpolation method to solve this restriction. 

Asad et al. proposed an n by K interpolation matrix $R$:

$$
R_{ij}:=
 \begin{cases} 
      1 & j=1, x_i \leq 1/K \\
      (j+1)-Kx_i & j = \lfloor K x_i \rfloor, x_i \in (1/K,1] \\
      Kx_i - (j-1) & j = \lceil Kx_i \rceil, x_i \in (1/K,1] \\
      0 & \text{otherwise},
   \end{cases}
$$
where $K = 2^{\lceil log_2 n \rceil}$.

Then we can instead solve the following optimization problem
$$
\hat{\beta} = \underset{\beta}{\mathrm{argmin}} \frac{1}{2}||y- RW^T\beta||_2^2 + \lambda||\beta||_1,
$$
where data points of $x_i$ is non-equispaced and $n$ is not necessarily a power of 2.

## implement the interpolation matrix 
```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

Here is a minimal example of interpolation matrix `R` when `n=7` and given non-equispaced data `x`. 
```{r}
n = 7
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
R = create_interpolation_matrix(x)
R
```

## simulation and performance

### example that works

We use the example from `SuSiE Trend Filtering` vignette. We simulate `n=1000` data points with `x` randomly drawn from a uniform distribution (0,1) i.e. `x` is not equally-spaced, and `y` has a piecewise-constant trend. Also notice that `n=1000` is not a power of two. The following graph shows the susie fit. 
```{r}
n = 1000
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,100),rep(1,100),rep(3,100),rep(-2,100),rep(0,600))
set.seed(1)
y = beta + rnorm(n, sd=0.1)
```

```{r}
W <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
R = create_interpolation_matrix(x)
# tcrossprod(R,W) computes R %*% t(W)
RW_t = tcrossprod(R, W)
```

```{r}
s = susie(RW_t, y, L = 30, estimate_prior_variance = TRUE)
```

```{r}
plot(x, y, pch=20, col="grey")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s), col='red', lwd=2)
```

Here is another example with more jumps.
```{r}
n1 = 1000
K1 = 2^(ceiling(log2(n1)))
set.seed(1)
x1 = sort(runif(n1, 0,1))
beta1 = c(rep(0,10),rep(1,20),rep(3,100),rep(-2,100),rep(0,50),rep(8,50),rep(3,50), rep(10,50), rep(0,(n-430)))
set.seed(1)
y1 = beta1 + rnorm(n1, sd=0.1)
```

```{r}
W1 <- t(GenW(n=K1, filter.number=1, family="DaubExPhase"))
R1 = create_interpolation_matrix(x1)
# tcrossprod(R,W) computes R %*% t(W)
RW_t1 = tcrossprod(R1, W1)
```

```{r}
s1 = susie(RW_t1, y1, L = 50, estimate_prior_variance = TRUE)
```

```{r}
plot(x1, y1, pch=20, col="grey")
lines(x1, beta1, col="black", lwd=2)
lines(x1, predict(s1), col='red', lwd=2)
```

### example that fails

I observe many bad fits when `n` is small and `y` has more than 2 jumps. The following is one of those examples. Here we simulate `n=150` data points with non-equispaced `x`. When fitting SuSiE, I have increased `L=80` but I still get an unsatisfactory fit. 

```{r}
n2 = 150
K2 = 2^(ceiling(log2(n2)))
set.seed(1)
x2 = sort(runif(n2, 0,1))
beta2 = c(rep(0,10),rep(1,20),rep(3,100),rep(-2,20))
set.seed(1)
y2 = beta2 + rnorm(n2, sd=0.1)
```

```{r}
W2 <- t(GenW(n=K2, filter.number=1, family="DaubExPhase"))
R2 = create_interpolation_matrix(x2)
# tcrossprod(R,W) computes R %*% t(W)
RW_t2 = tcrossprod(R2, W2)
```

```{r}
s2 = susie(RW_t2, y2, L = 80, estimate_prior_variance = TRUE)
```

```{r}
plot(x2, y2, pch=20, col="grey")
lines(x2, beta2, col="black", lwd=2)
lines(x2, predict(s2), col='red', lwd=2)
```


