---
title: "Wavelet-based SuSiE Analysis 20190228-2"
author: "Kaiqian Zhang"
date: "2/28/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this vignette, we investigate two interpolation schemes. We have different K's in two schemes.

* Scheme 1: we set 

$$
n = \text{length}(x) \\
K = 2^{\lceil \text{log}_2n \rceil}
$$

* Scheme 2: we set

$$
m = \lceil 10\%\text{quantiele}(|x_i-x_{i+1}|) \rceil \\
K = 2^{\lceil \text{log}_2m \rceil}.
$$

## Results

* Scheme 2 solves my bad fit in the simulation example 2 as I mentioned in the previous analysis.

* Scheme 2 also works when my data have large gaps. Please see simulation example 3. 

* We can still find very bad fits for using scheme 2. I think this is due to local optimum problem. Please see simulation example 4. I notice that when my data are evenly spaced, I can get a relatively good fit. So I intialize from there to fit my example 4 using scheme 1. The reason I used scheme 1 is that it has the same K as evenly spaced data. I now can get a good fit for example 4. So I hypothesize that SuSiE using scheme 2 is also having a local optimum problem. But I could not find a decent initialization to confirm my hypothesis. 

## Set up
```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
```

Function for interpolation scheme 1:
```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix_original = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

Function for interpolation scheme 2:
```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  #m = 10% * abs(x_i - x_{i+1})
  m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
  K = 2^(ceiling(log2(m)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

Function for creating Wt
```{r}
#' @param K is the dimension of decinmated W
#' @return a K by K matrix, which the transpose of decimated Harr wavelet transform matrix 
create_Wt_DWT = function(K){
  return(GenW(n=K, filter.number=1, family="DaubExPhase"))
}
```

Function for simulate y
```{r}
create_y = function(x, beta.function, n, sd){
  set.seed(1)
  return(beta.function(x) + rnorm(n, sd=sd))
}
```

Function for creating gaps in x
```{r}
create_gaps_x = function(n){
  set.seed(1)
  x = sort(runif(n, 0, 1))
  cutidx1 = which(x>0.2 & x<0.3)
  cutidx2 = which(x>0.6 & x<0.8)
  x.cut = x[-c(cutidx1, cutidx2)]
  set.seed(1)
  x.add1 = runif(length(cutidx1), 0.8, 0.9)
  x.add2 = runif(length(cutidx2), 0.3, 0.4)
  x.res = sort(c(x.cut, x.add1, x.add2))
}
```

## Simulation example 2 using interpolation scheme 2 

Back to simulation example 2, by using interpolation scheme 2, we can have a better fit. Red line is the new fit using scheme 2, and the blue line is the previous version. 

```{r}
f.beta4 = function(x){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < 0.6){
      beta[i] = -3
    } else if(x[i] >= 0.6 & x[i] < 0.65){
      beta[i] = 7
    } else if(x[i] >= 0.65 & x[i] < 0.7){
      beta[i] = 0
    } else {
      beta[i] = 2
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta4, n, sd)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
```

```{r}
#Create RWt
R.previous = create_interpolation_matrix_original(x)
Wt.DWT.previous = create_Wt_DWT(2^(ceiling(log2(n))))
RWt.previous = R.previous %*% Wt.DWT.previous
#Run SuSiE
s.previous = susie(RWt.previous, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
#Run SuSiE
s = susie(RWt, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 1: fit for simulation example 2")
lines(x, f.beta4(x), col="black", lwd=2)
lines(x, predict(s.previous), col="blue", lwd=2)
lines(x, predict(s), col="red", lwd=2)
```

## Simulation example 3

Our new interpolation scheme also works when there are large gaps in the data.

```{r}
sd = 0.1
n = 100
set.seed(1)
x = create_gaps_x(n)
y = create_y(x, beta.function=f.beta4, n, sd)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
```

```{r}
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
#Run SuSiE
s = susie(RWt, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 2: example with two large gaps in xs")
lines(x, f.beta4(x), col="black", lwd=2)
lines(x, predict(s), col="red", lwd=2)
```

## Simulation example 4

We observe many good fits in our simulations using the new interpolation scheme but we observe the following example that gives us a really bad fit. And I hypothesize that this is due to local optimum problem in SuSiE. Here is the example.

```{r}
f.beta5 = function(x){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < 0.1){
      beta[i] = -3
    } else if(x[i] >= 0.1 & x[i] < 0.6){
      beta[i] = 5
    } else if(x[i] >= 0.6 & x[i] < 0.7){
      beta[i] = -1
    } else if(x[i] >= 0.7 & x[i] < 0.9){
      beta[i] = 3
    }else{
      beta[i] = -3
    }
  }
  return(beta)
}
```

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta5, n, sd)
m = as.numeric(ceiling(1/quantile(sort(abs(diff(x))),0.1)))
K = 2^(ceiling(log2(m)))
```

```{r}
#Create RWt
R = create_interpolation_matrix(x)
Wt.DWT = create_Wt_DWT(K)
RWt = R %*% Wt.DWT
#Run SuSiE
s = susie(RWt, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 3: fit for simulation example 4")
lines(x, f.beta5(x), col="black", lwd=2)
lines(x, predict(s), col="red", lwd=2)
```

I notice that when my data are evenly spaced, I can get a relatively good fit.

```{r}
sd = 0.1
n = 100
set.seed(1)
x = seq(n)/n
y = create_y(x, beta.function=f.beta5, n, sd)
```

```{r}
#Create RWt
R.original = create_interpolation_matrix_original(x)
Wt.DWT.original = create_Wt_DWT(2^(ceiling(log2(n))))
RWt.original = R.original %*% Wt.DWT.original
```

```{r}
#Run SuSiE
s.original.even = susie(RWt.original, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 4: fit for evenly spaced data")
lines(x, f.beta5(x), col="black", lwd=2)
lines(x, predict(s.original.even), col="blue", lwd=2)
```

Then we initialize from here and fit the simulation example 4. We obtain a nice fit after initialization. We cannot perform this initialization on the new interpolation scheme since K is different in the evenly spaced and irregularly spaced data. 

```{r}
sd = 0.1
n = 100
set.seed(1)
x = sort(runif(n, 0, 1))
y = create_y(x, beta.function=f.beta5, n, sd)
```

```{r}
#Create RWt
R.original = create_interpolation_matrix_original(x)
Wt.DWT.original = create_Wt_DWT(2^(ceiling(log2(n))))
RWt.original = R.original %*% Wt.DWT.original
```

```{r}
#Run SuSiE
s.original.uneven = susie(RWt.original, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim', s_init = s.original.even)
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 5: initialize from the evenly-spaced fit")
lines(x, f.beta5(x), col="black", lwd=2)
lines(x, predict(s.original.uneven), col="green", lwd=2)
```











