---
title: "Wavelet-based SuSiE Analysis 20190222-2"
author: "Kaiqian Zhang"
date: "2/22/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We create our interpolation matrix based on the our observed data `x`, which is irregularly spaced. We, however, always observe some artifacts in the SuSiE fit. We notice that if we create the interpolation matrix from corresponding evenly spaced `x`, we obtain a really good fit. Our data here are irregularly spaced and not a power of 2 (n=100). Please see the figure below. The blue line is using interpolation matrix created from real irregularly spaced `x`. The green line is using interpolation matrix created from evenly spaced `x`. The green one renders a better fit. 
```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
library(smashr)
library(varbvs)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x1 = sort(runif(n, 0,1))
x2 = seq(1,n)/n
#beta = c(rep(0,20), rep(3,20), rep(-3,20), rep(5,40))
beta = c(rep(0,10), rep(3,30), rep(-7,30), rep(10,30))
#beta = c(rep(0,30), rep(7,2), rep(0,100-32))
set.seed(1)
y = beta + rnorm(n, sd=0.2)
```

```{r}
R.noneven = create_interpolation_matrix(x1)
R.even = create_interpolation_matrix(x2)
```

```{r}
W.decimated = t(GenW(n=K, filter.number=1, family="DaubExPhase"))
# tcrossprod(R,W) computes R %*% t(W)
RW_t.noneven = tcrossprod(R.noneven, W.decimated)
RW_t.even = tcrossprod(R.even, W.decimated)
```

```{r}
s.susie.noneven = susie(RW_t.noneven, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
s.susie.even = susie(RW_t.even, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x1, y, pch=20, col="grey", main = "Figure 2")
lines(x1, beta, col="black", lwd=2)
lines(x1, predict(s.susie.noneven), col="blue", lwd=2)
lines(x1, predict(s.susie.even), col="green", lwd=2)
```
