---
title: "Wavelet-based SuSiE Analysis 20190129"
author: "Kaiqian Zhang"
date: "1/29/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
library(glmnet)
library(smashr)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

## examples with "bad" fit

Last time we gave an example where the performance is not very satisfactory. This time we want to find worse examples and try to understand why those examples fail. Last time, we notice that when `n` is small and you have more than one jump in the data, the fit would not be decent. Figure 1 and Figure 2 both use `n=100` (not a power of 2) non-equispaced data points. We observe that when we have more than one jump, SuSiE provides a bad fit.  
```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,20), rep(3,80))
set.seed(1)
y = beta + rnorm(n, sd=0.1)
```

```{r}
W <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
R = create_interpolation_matrix(x)
# tcrossprod(R,W) computes R %*% t(W)
RW_t = tcrossprod(R, W)
```

```{r}
s = susie(RW_t, y, L = 50, estimate_prior_variance = TRUE)
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 1: non-equispaced 100 data points fit for one jump")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s), col='red', lwd=2)
```

```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,20), rep(3,20), rep(-3,20), rep(5,40))
set.seed(1)
y = beta + rnorm(n, sd=0.1)
```

```{r}
W <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
R = create_interpolation_matrix(x)
# tcrossprod(R,W) computes R %*% t(W)
RW_t = tcrossprod(R, W)
```

```{r}
s = susie(RW_t, y, L = 50, estimate_prior_variance = TRUE)
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 2: non-equispaced 100 data points fit for three jumps")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s), col='red', lwd=2)
```

Putting aside non-equispaced and not a power of 2 implementation, we first want to understand the following bad fit. Suppose we have `n=128` (i.e. n is a power of 2) equal-spaced data points with three jumps. Fitting with orthogoal wavelet basis, we still get a bad fit. We can observe this from Figure 3 below. 
```{r}
n = 128
beta = c(rep(0,20), rep(3,20), rep(-3,20), rep(5,68))
set.seed(1)
y = beta + rnorm(n, sd=0.1)
```

```{r}
W_t <- GenW(n=n, filter.number=1, family="DaubExPhase")
```

```{r}
s = susie(W_t, y, L = 100, estimate_prior_variance = TRUE)
```

```{r}
plot(y, pch=20, col="grey", main="Figure 3: equal-space 128 data points with three jumps")
lines(beta, col="black", lwd=2)
lines(predict(s), col='red', lwd=2)
```

We apply wavelet decomposition of truth and data and plot their coefficients. We notice that both decompositions identify three non-zeros around same locations. And the number of non-zeros indicates that beta is pretty sparse. Our question becomes if coefficients are the same, is there anything wrong with prediction?
```{r}
s_true = susie(W_t, beta, L=50, estimate_prior_variance = TRUE)
```

```{r}
plot(coef(s_true), main='wavelet decomposition of truth')
```

```{r}
plot(coef(s), main='wavelet decomposition of data')
```























