---
title: "Wavelet-based SuSiE Analysis 20190215"
author: "Kaiqian Zhang"
date: "2/15/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Previously, we applied L0 trend filtering to SuSiE for solving change point problem. This method, however, failed in the case when two change points in a quick succession. Since columns of `X` constructed by L0 trend filtering are correlated, two effects from the "hard" case cancel with each other. 

We further used wavelets to construct `X` since wavelets are orthogonal. One pain for using wavelets is that it requires the data to be equal-spaced and the number to be a power of two. We used interpolation matrix introduced in the paper [(paper link)](https://papers.nips.cc/paper/8112-wavelet-regression-and-additive-models-for-irregularly-spaced-data.pdf) to adjust this problem. But due to issues with estimating prior variance about SuSiE itself, we initially did not get a good fit. After solving the bug, we now try our previous example using `optimV_method="optim"` and `optimV_method="EM"`.

## Results

We show three examples below. Notice that all data generated are non-equal spaced and the number is not a power of two `n=100`.

**Summary**

- Figure 1 shows susie fit for non-equal spaced `n=100` data points. Red line uses `optim` and blue line uses `EM`. We observe that using `optim` would give us a much better fit. So we use `optim` in the other two examples afterwards. We notice that when `n=100`, the fit from red line is still not perfect. 

- Figure 2 shows another example using non-equal spaced `n=100` data points. The fit is okay but not good enough. 

- Figure 3 shows the "hard" case mentioned in the change point problem. Notice that our data are non-equal spaced and we have `n=100` data points. It seems like using `interpolation matrix + wavelet matrix + optim`, we can solve the "hard" change point problem. 


```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
library(glmnet)
library(smashr)
library(MASS)
library(L0Learn)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,20), rep(3,20), rep(-3,20), rep(5,40))
set.seed(1)
y = beta + rnorm(n, sd=0.2)
```

```{r}
W <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
R = create_interpolation_matrix(x)
# tcrossprod(R,W) computes R %*% t(W)
RW_t = tcrossprod(R, W)
```

```{r}
s.em = susie(RW_t, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'EM')
```

```{r}
s.optim = susie(RW_t, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

Objectives obtained by using `EM` and `optim`:
```{r}
susie_get_objective(s.em)
susie_get_objective(s.optim)
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 1")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s.optim), col="red", lwd=2)
lines(x, predict(s.em), col="blue", lwd=2)
```

```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,10), rep(3,30), rep(-7,30), rep(10,30))
set.seed(1)
y = beta + rnorm(n, sd=0.1)
```

```{r}
W <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
R = create_interpolation_matrix(x)
# tcrossprod(R,W) computes R %*% t(W)
RW_t = tcrossprod(R, W)
```

```{r}
s.optim = susie(RW_t, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 2")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s.optim), col='red', lwd=2)
```


```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,30), rep(7,2), rep(0,100-32))
set.seed(1)
y = beta + rnorm(n, sd=0.2)
```

```{r}
W <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
R = create_interpolation_matrix(x)
# tcrossprod(R,W) computes R %*% t(W)
RW_t = tcrossprod(R, W)
```

```{r}
s.optim = susie(RW_t, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

```{r}
plot(x, y, pch=20, col="grey", main = "Figure 3")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s.optim), col='red', lwd=2)
```












