---
title: "Wavelet-based SuSiE Analysis 20190121"
author: "Kaiqian Zhang"
date: "1/21/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
library(glmnet)
```

Previously, we simulated `n=128` piecewise-constant data points and apply both DWT and UDWT methods with prior estimated. There is not much difference between two fits. So I asked a question:

* what is advantage we expect that UDWT could bring over DWT?

Suggested by Matthew, I read the paper `Translation-Invariant De-Noising` [link](https://web.stanford.edu/dept/statistics/cgi-bin/donoho/wp-content/uploads/2018/08/TIDeNoise.pdf) and learned that:

* Using UDWT can reduce visual artifacts. For example, in the wavelet-denoising case, the artifacts have to do with behavior near singularities. In the neighborhood of discontinuities, wavelet de-noising can exhibit pusedo-Gibbs phenomena.

* Gibbs phenomena is something would happen in the Fourier-based de-noising: you will observe upshoot and downshoot of a specific target level at the discontinuity of jumps. 

* Importantly, the size of those artifacts is connected intimately with the actual location of the discontinuity. For example, when using Haar wavelets, a discontinuity precisely at location `n/2` will lead to essentially no pseudo-Gibbs oscillations;
a discontinuity near a binary irrational like `n/3` will lead to significant pseudo-Gibbs oscillations. 

## Pseudo-Gibbs phenomena and discontinuities

Based on my reading, I want to investigate the relationship between pseudo-Gibbs phenomena and the location of discontinuities. I expect to see that in the DWT, I would not see visual artifacts if the discontinuity happens at `n/2`. And I would observe visual artifacts if the discontinuity is at `n/3`. 

We observe from the following two figures `DWT for fitting data with discontinuity at n/2` and `DWT for fitting data with discontinuity at n/3` that:

* SuSiE fits really well when the discontinuity happens at n/2.

* We do have visual artifacts when the jump is not at n/2.

```{r}
n = 64
beta_jump2 = c(rep(0,32), rep(5,32))
beta_jump3 = c(rep(0,21), rep(5,43))
set.seed(1)
y_jump2 = beta_jump2 + rnorm(n, sd=0.1)
set.seed(1)
y_jump3 = beta_jump3 + rnorm(n, sd=0.1)
```

```{r}
Haar_matrix <- t(GenW(n=n, filter.number=1, family="DaubExPhase"))
DWT_W_T <- t(Haar_matrix)
```

```{r}
s_jump2_DWT = susie(DWT_W_T, y_jump2, estimate_prior_variance = TRUE)
```

```{r}
plot(y_jump2,pch=20, col="grey", main="DWT for fitting data with discontinuity at n/2")
lines(beta_jump2,col="black",lwd=2)
lines(predict(s_jump2_DWT), col='red', lwd=2)
```

```{r}
s_jump3_DWT = susie(DWT_W_T, y_jump3, estimate_prior_variance = TRUE)
```


```{r}
plot(y_jump3,pch=20, col="grey", main="DWT for fitting data with discontinuity at n/3")
lines(beta_jump3,col="black",lwd=2)
lines(predict(s_jump3_DWT), col='red', lwd=2)
```

## What about UDWT?
I want to further look at how UDWT behaves. I expect that UDWT can solve the visual artifacts. But the following figure is not what I expected. My hypothesis: my UDWT matrix is wrong?

```{r}
n_Haar_matrix_inv <- DWT_W_T
for(i in 1:(n-1)){
  order <- shift(seq(n),i)
  Haar_matrix_shifted_inv <- t(Haar_matrix[,order]) 
  n_Haar_matrix_inv <- cbind(n_Haar_matrix_inv, Haar_matrix_shifted_inv) #Each column is a basis for sparse regression
}
duplicated_columns <- duplicated(t(n_Haar_matrix_inv))
UDWT_W_T <- n_Haar_matrix_inv[, !duplicated_columns]
```

```{r}
s_jump3_UDWT = susie(UDWT_W_T, y_jump3, estimate_prior_variance = TRUE)
```

```{r}
plot(y_jump3,pch=20, col="grey", main="UDWT for fitting data with discontinuity at n/3")
lines(beta_jump3,col="black",lwd=2)
lines(predict(s_jump3_UDWT), col='blue', lwd=2)
```

## Other sparse regression methods

Is it really the problem of my undecimated wavelet matrix `UDWT_W_T` or it is a problem with SuSiE? So I tried other sparse regression methods. The fit using UDWT bases and glmnet is still bad. So something is wrong with my `UDWT_W_T` matrix. 

```{r}
s_glmnetcv = cv.glmnet(UDWT_W_T, y_jump3, type.measure = "mse", nfolds = 20)
pred_glmnetcv = predict(s_glmnetcv, newx = UDWT_W_T, s = "lambda.min")
```

```{r}
plot(y_jump3,pch=20, col="grey", main="glmnet-UDWT for fitting data with discontinuity at n/3")
lines(beta_jump3,col="black",lwd=2)
lines(pred_glmnetcv, col='orange', lwd=2)
```

```{r}
s_glmnetcv_DWT = cv.glmnet(DWT_W_T, y_jump3, type.measure = "mse", nfolds = 20)
pred_glmnetcv_DWT = predict(s_glmnetcv_DWT, newx = DWT_W_T, s = "lambda.min")
```

```{r}
plot(y_jump3,pch=20, col="grey", main="glmnet-DWT for fitting data with discontinuity at n/3")
lines(beta_jump3,col="black",lwd=2)
lines(pred_glmnetcv_DWT, col='orange', lwd=2)
```

## Fix UDWT matrix
Let's start from something simple and try to debug here. Set `n=8`. 
```{r}
n = 8
Haar_matrix <- t(GenW(n=n, filter.number=1, family="DaubExPhase"))
DWT_W_T <- t(Haar_matrix)
n_Haar_matrix_inv <- DWT_W_T
for(i in 1:(n-1)){
  order <- shift(seq(n),i)
  Haar_matrix_shifted_inv <- t(Haar_matrix[,order]) 
  n_Haar_matrix_inv <- cbind(n_Haar_matrix_inv, Haar_matrix_shifted_inv) #Each column is a basis for sparse regression
}
duplicated_columns <- duplicated(t(n_Haar_matrix_inv))
UDWT_W_T <- n_Haar_matrix_inv[, !duplicated_columns]
```

```{r}
Haar_matrix[Haar_matrix>0] = 1
Haar_matrix[Haar_matrix<0] = -1
Haar_matrix
```

```{r}
UDWT_W_T[UDWT_W_T>0] = 1
UDWT_W_T[UDWT_W_T<0] = -1
UDWT_W_T
```













