---
title: "Wavelet-based SuSiE Analysis 20190223"
author: "Kaiqian Zhang"
date: "2/23/2019"
header-includes:
   - \usepackage{amsmath}
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

Our goal is to apply wavelets in SuSiE to better solve nonparametric regression problems. We start from the simple change point problem and try to use Harr wavelet here. This idea works pretty well given evenly spaced data and the number is a power of 2. We further want to extend `wavelet-based SuSiE` to allow for irregularly spaced data and the number is not a power of 2.

## Previously...

Given observed data $(x_i, y_i)$ for $i = 1, \dots, i$. We create an interpolation matrix `R` by using observed data `x_i`'s. This idea comes from Asad's paper in 2018. So our framework becomes "interpolation R + discrete wavelet transform W + SuSiE". 

Previously, we applied this framework on our simulated piecewise constant data. But we always got visual artifacts in the SuSiE fit. This is due to our simulated data `y` did not depend on `x`. After correcting our simulation, we notice that this framework has a really decent performance and certainly solves the problem of irregularly spaced data and not a power of 2. 

## Simulation

We simulated data `y` whose underlying trend is piecewise constant and this trend depends on `x`'s. Let `\beta` be the underlying trend of `y`. So we have, for $i = 1, \dots, n$,

$$
\beta_i = 
\begin{cases} 
      -2 & x_i < 0.25 \\
      7 & 0.25\leq x_i < 0.5 \\
      0 & 0.5\leq x_i < 0.85 \\
      3 & 0.85 \leq x_i \leq 1.
\end{cases}
$$
Our `x`'s are irregularly spaced and we intentionally create large gaps in the `x`'s. The number of data is `n=100`. 

## Result

From this minimal simulation example, we notice that "interpolation + DWT + SuSiE" has a nice performance for irregularly spaced data and the number is not a power of 2! Please see the figure below. The black line is the underlying trend and the blue line is the SuSiE fit. The blue line is on top of the black line perfectly. 

## Code details

Install packages.
```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
```

Create interpolation matrix function:
```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

Underlying trend beta function:
```{r}
f.beta = function(x){
  n = length(x)
  beta = numeric(n)
  for (i in 1:n){
    if (x[i] < 0.25){
      beta[i] = -2
    } else if(x[i] >= 0.25 & x[i] < 0.5){
      beta[i] = 7
    } else if(x[i] >= 0.5 & x[i] < 0.85){
      beta[i] = 0
    } else{
      beta[i] = 3
    }
  }
  return(beta)
}
```

Simulate irregularly spaced `x` and create several large gaps in `x`'s. 
```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))

cutidx1 = which(x<0.3 & x>0.1)
cutidx2 = which(x<0.7 & x>0.6)
cutidx3 = which(x<0.9 & x>0.8)
x1 = x[-c(cutidx1, cutidx2, cutidx3)]
set.seed(1)
new1 = runif(length(cutidx1), 0, 0.1)
set.seed(1)
new2 = runif(length(cutidx2), 0.3,0.6)
set.seed(1)
new3 = runif(length(cutidx3), 0.7, 0.8)
x.noneven = sort(c(x1, new1, new2, new3))
```

Simulate `y` based on `x`.
```{r}
set.seed(1)
y = f.beta(x.noneven) + rnorm(n, sd=0.2)
```

Create interpolation matrix `R` and discrete wavelet transform matrix `W`.
```{r}
R = create_interpolation_matrix(x.noneven)
W = t(GenW(n=K, filter.number=1, family="DaubExPhase"))
RW_t = tcrossprod(R,W)
```

Run SuSiE.
```{r}
s = susie(RW_t, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

Plot the SuSiE fit.
```{r}
plot(x.noneven, y, pch=20, col="grey", main = "Figure 1: fit for interpolation + DWT + SuSiE")
lines(x.noneven, f.beta(x.noneven), col="black", lwd=2)
lines(x.noneven, predict(s), col="blue", lwd=2)
```










