---
title: "Wavelet-based SuSiE Analysis 20190217"
author: "Kaiqian Zhang"
date: "2/17/2019"
output: 
  workflowr::wflow_html:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal 

Previously, we've tried "interpolation + DWT + SuSiE". Now, we want to investigate whether "undecimated" wavelet transform would work after interpolation, i.e. "interpolation + UDWT + SuSiE". 

## Code details
```{r, warning=FALSE, include=FALSE}
library(wavethresh)
library(binhf) #Package for shift function
library(susieR)
library(smashr)
```

```{r}
#' @param x is an n-vector of data
#' @return R an n by K interpolation matrix
create_interpolation_matrix = function(x){
  n = length(x)
  K = 2^(ceiling(log2(n)))
  R = matrix(0, n, K)
  for (i in 1:n){
    for (j in 1:K){
      if (j == 1 & x[i] <= 1/K){
        R[i,j] = 1
      } else if (j == floor(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = (j+1) - K*x[i]
      } else if (j == ceiling(K*x[i]) & x[i] > 1/K & x[i] <=1){
        R[i,j] = K*x[i] - (j-1)
      } else R[i,j] = 0
    }
  }
  return(R)
}
```

```{r}
n = 100
K = 2^(ceiling(log2(n)))
set.seed(1)
x = sort(runif(n, 0,1))
beta = c(rep(0,20), rep(3,20), rep(-3,20), rep(5,40))
#beta = c(rep(0,10), rep(3,30), rep(-7,30), rep(10,30))
#beta = c(rep(0,30), rep(7,2), rep(0,100-32))
set.seed(1)
y = beta + rnorm(n, sd=0.2)
```

Create an interpolation R matrix. R is an n by K matrix, where K=2^J.
```{r}
R = create_interpolation_matrix(x)
```

Create a decimated W matrix. W.decimated is a K by K matrix. 
```{r}
W.decimated <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
# tcrossprod(R,W) computes R %*% t(W)
RW_t.decimated = tcrossprod(R, W.decimated)
zero.idx = which(colSums(RW_t.decimated)==0)
RW_t.decimated = RW_t.decimated[,-zero.idx]
```

Create an undecimated W matrix. W.undecimated is a K by Klog2(K) matrix.
```{r}
Haar_matrix <- t(GenW(n=K, filter.number=1, family="DaubExPhase"))
DWT_W_T <- t(Haar_matrix)
n_Haar_matrix_inv <- DWT_W_T
for(i in 1:(K-1)){
  order <- shift(seq(K),i)
  Haar_matrix_shifted_inv <- t(Haar_matrix[,order])
  n_Haar_matrix_inv <- cbind(n_Haar_matrix_inv, Haar_matrix_shifted_inv) #Each column is a basis for sparse regression
}
duplicated_columns <- duplicated(t(n_Haar_matrix_inv))
UDWT_W_T <- n_Haar_matrix_inv[, !duplicated_columns]
```

```{r}
RW_t.undecimated = R %*% UDWT_W_T
zero.idx = which(colSums(RW_t.undecimated)==0)
RW_t.undecimated = RW_t.undecimated[,-zero.idx]
```

SuSiE with decimated wavelet basis
```{r}
s.decimated = susie(RW_t.decimated, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

SuSiE with undecimated wavelet basis
```{r}
s.undecimated = susie(RW_t.undecimated, y, L=50, estimate_prior_variance = TRUE, optimV_method = 'optim')
```

The red line shows DWT-based SuSiE fit and blue one depicts UDWT version. Undecimated version does not render a better fit as we expected... 
```{r}
plot(x, y, pch=20, col="grey", main = "DWT and UDWT-based SuSiE fit")
lines(x, beta, col="black", lwd=2)
lines(x, predict(s.decimated), col='red', lwd=2)
lines(x, predict(s.undecimated), col='blue', lwd=2)
```


